---
title: "Class Demo"
author: "Sujay"
date: "2022-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Yo what up g
```{r}
x <- 5
y <- 3
x+y

```


Lets explore Youth Risk Behavior Surveillance System Survey
(YRBSSS)

```{r}
library(openintro)
library(ggplot2)

ggplot(data = yrbss) +
  geom_bar(aes(x = school_night_hours_sleep))



```


Stats inference question 1:
Are 12yr old children in the US physically active?
Suppose we want to test the claim that 12yr old children are active twice a week
on average


Prepare: Choosing a test statistic (mean of physically active data and compare to mean)
calculate mean

x-bar = 3.875

Check: Survey respondents are independent
The responses, despite being self-reported, are "true"
We dont know the exact sampling method but there is not way our 26 12yr olds are representative of 12 yr olds in the US


Calculate:

Conclude:


```{r}
twelve <- subset(yrbss, age == 12)
mean(twelve$physically_active_7d, na.rm = TRUE)

table(yrbss$age)
```



Lets look at some of our oxymoronic children
Lets find all of the students who text and drive every day    *AND*
Lets find all of the students who always wear a helmet while riding a bicycle

```{r}
oxymorons <- subset(yrbss, text_while_driving_30d == 30 & helmet_12m == "always")
```

Lets test the claim that these students are physically active, on average 2 days per week

Prepare: My test statistic, the sample average is

```{r}
mean(oxymorons$physically_active_7d, na.rm = TRUE)
```

Check: Assumptions are met

Calculate:  
```{r}
sd(oxymorons$physically_active_7d, na.rm = TRUE)
```



t-test

claim:
mean = 2

x_bar11 = 4.09

s = 3.11


```{r}
oxymorons$physically_active_7d

#Lets remove missing values
sd(oxymorons$physically_active_7d, na.rm = TRUE)

noMissing <- na.omit(oxymorons$physically_active_7d)

sd(noMissing)

```

Lets treat these 11 values as my population
sigma = 3.11
```{r}
sd(sample(noMissing, size = 3))

```


Anytime I have to estimate my population standard deviation (sigma)
using my sample standard deviation (s), the x-bar actually follows a t-distribution
instead of a normal distribution

```{r}
#Find middle 95% of values from N(0,1)
qnorm(.025)
qnorm(.975)
```
2.5% of my data are passed 2 standard deviations on either side



Now how does this work for a t-distribution?

Suppose I have n = 3
```{r}
n <- 3
qt(.025, n-1)

n <- 100
qt(.025, n-1)
```


Suppose I want to find my middle 99% of values
```{r}
qnorm(.005)


#quantile()
quantile(na.omit(yrbss$height), .05)
#works with actual data not theoretical like other two
```



Lets do a t-test

Prepare: 
claim is than mu = 2, x-bar_11 = 4.09, alpha = .05

check:
data are independent and randomly sampled
(we'll assume our missing values are "missing at random")


calculate:
```{r}
#Lets start with a 95% "confidence interval" - (middle 95% of values in t-distribution)
qt(.025, 10)
```

Standard Error = s/sqrt(n)
```{r}
se <- sd(noMissing)/sqrt(11)

t <- qt(.025, 10)
t*se

#t-statistic = t-value *s/sqrt(11)
```

95% of our data is between -.09 and 4.09, so claim is reasonable

But, -.09 is not possible for activity days

```{r}
#Lets calculate a p-value
1 - pt(2.23, 10)

```



How do I do a t-test() function to do all of this automatically?
```{r}
t.test(noMissing, mu = 2, conf.level = .95)
```

confidence interval is centered at 4.09 and seeing if 2 is in the interval 
which is the same as centering at 2 and seeing if 4.09 is in interval



Lets run the fastest t-test ever

test the claim that the average height of a high schooler is 1.72m

```{r}
t.test(yrbss$height, mu = 1.72)
```
the t = -30 means it is 30 standard deviations away
p value so low because the standard deviation is so small because sample size so big




